---
title: "Data Engineering Project 2 - Campaign Contribution Analysis"
author: "Austin Smith"
date: "11/28/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

```{r}
# read in the data and combine the two tables 
contribution_data <- read.csv('contribution_data.csv')

# remove the index column and the CID column, both of which are not needed
contribution_data <- contribution_data[ -c(1,6) ]

head(contribution_data)
```

## Part 2: Analysis
(please see Jupyter Notebook "Part 1 - Campaign Contributions Data Pull" for explanation of how the data was obtained.)

This data set, which includes all of the congress people, their party, state, and the amounts they received from all of their major contributors, can be mined for all sorts of interesting information. 

In the code below, I perform analysis to collect the following: 

- Top 10 contributors across both political parties
- Top 10 contributors by party
- The 10 congress people who received the most in total contributions

### Top 10 Contributors Across all Parties
```{r cars}
# top 10 contributors across both parties
top10_contributors <- contribution_data %>%
  select(contributor,total_amount) %>%
  group_by(contributor) %>%
  summarize(sum(total_amount)) 

# for whatever reason, I can't get this to stop throwing an error unless I make it a new df, so doing so here. 
top10_contribs <- rename(top10_contributors, total_contributions = 'sum(total_amount)')

top10_contribs %>%
  arrange(desc(total_contributions)) %>%
  head(10)
```

Democracy Engine contributed the most to congress people in the most recent election cycle (which, note, covers the last campaign for every congress person, be that on a 2-year or 6-year cycle). A quick search on Open Secrets of Democracy Engine informed me that it is a democratic supporting organization; its top "contributees" in 2020 were Joe Biden, Mark Kelly, and Steve Bullock. Interestingly, Democracy Engine is just a donor platform; it, in itself, is non-partisan. Its platform allows donors from any political party to contribute to any candidate. It just so happens that the majority of their users supported democrats during the last election. 

Of the other top 5 contributors; Alphabet, parent company of Google, gave entirely to democratic candidates in 2020; Votesane PAC contributed solely to republican candidates; and both universities and Microsoft Corp gave entirely to democratic candidates (and Bernie). Among the major donors, democratic supporters were the largest contributors in recent elections. 

### Top 10 Contributors to Republican Party 
```{r}
top10_repubs <- contribution_data %>%
  filter(party == 'Republican') %>%
  select(contributor,party,total_amount) %>%
  group_by(contributor) %>%
  summarize(sum(total_amount)) 

top10_repubs <- rename(top10_repubs, total_contributions = 'sum(total_amount)')

top10_repubs %>%
  arrange(desc(total_contributions)) %>%
  head(10)
```

It makes sense that the Blackstone Group, a large investment firm, and Raytheon, a major defense contractor, would be large republican supporters. It's interesting to see a couple of major banks and Boeing Co. in the top 10, as well. Less regulation is beneficial in those industries, so their support of republican candidates is logical.

The Republican Jewish Coalition's support is evident. The Club for Growth is a PAC focused on lowering taxes and government regulation. NorPAC, interestingly, is non-partisan; its main goal is to improve US-Israel relations. Its support between republican and democratic party candidates has been split over the years. 

### Top 10 Contributors to Democratic Party
```{r}
top10_dems <- contribution_data %>%
  filter(party == 'Democrat') %>%
  select(contributor,party,total_amount) %>%
  group_by(contributor) %>%
  summarize(sum(total_amount)) 

top10_dems <- rename(top10_dems, total_contributions = 'sum(total_amount)')

top10_dems %>%
  arrange(desc(total_contributions)) %>%
  head(10)
```

We have already noted the top five here, as they were among the top overall contributors. It's interesting to see Blue Cross/Blue Shield shows up in the top 10 as I would have guessed an insurance company might have leaned more red. A couple of other universities crack the top 10, as does a pro-labor rights group and Emily's List, a PAC that aims to elect female democrats who are in favor of abortion rights.  

###  Top 10 Most Well-Funded Candidates
```{r}
# top 10 funded candidates of last election cycle
top10_members <- contribution_data %>%
  select(politician, party,state,district,total_amount) %>%
  group_by(politician) %>%
  mutate(total_contributed = sum(total_amount)) %>%
  ungroup() %>%
  select(politician,party,state,district,total_contributed) %>%
  distinct() %>%
  arrange(desc(total_contributed)) %>%
  head(10)

top10_members
```

All of the top 10 best funded congress people in their last election cycles (respectively, as not all of them were up for election in the last cycle) were Senators. Democrats landed in the top 2 places, and claimed 6 of the top 10 spots overall. 

Senator candidates from Arizona and Georgia, both key swing states in the last election, received substantial contributions, as did very well-known and influential senators like Mitch McConnell and Lindsey Graham.